---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

```{r}
library(dplyr)
library(ggplot2)
library(lattice)
```

# Loading and preprocessing the data
```{r}
download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip", destfile="activity.zip", method="curl")
unzip(zipfile = "activity.zip")
```

```{r}
data <- read.csv("activity.csv", stringsAsFactors=FALSE)
data$date <- as.Date(data$date, "%Y-%m-%d")
```

# What is mean total number of steps taken per day?
```{r}
daily <- data[,c("steps", "date")]
daily <- aggregate(steps ~ date, daily, FUN=mean)
hist(daily$steps, xlab="Steps", main = "Histogram of number of steps per day")
```

## Mean value
```{r}
mean(daily$steps, na.rm = TRUE)
```

## Median value
```{r}
median(daily$steps, na.rm = TRUE)
```

# What is the average daily activity pattern?
```{r}
byInterval <- data[,c("steps", "interval")]
byInterval <- aggregate(steps ~ interval, byInterval, FUN=mean)
plot(byInterval, type="l")
```

## Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
maxRow <- which.max(byInterval$steps)
byInterval[maxRow,"interval"]
```

# Imputing missing values
## Number of missing values
```{r}
sum(!complete.cases(data))
```

## Impute the values
Basically you can put anything into my_impute function. 
For illustration purposes I've picked the simplest possible thing - a constant :)
```{r}
my_impute <- function(date, interval) {
  return(36);
}
imputed <- data
imputed$steps <- ifelse(is.na(data$steps), my_impute(imputed$date, imputed$interval), data$steps)
```

## Let's see how the statistics changed
```{r}
imputedDaily <- imputed[,c("steps", "date")]
imputedDaily <- aggregate(steps ~ date, imputedDaily, FUN=mean)
hist(imputedDaily$steps)
```

### Mean value
```{r}
mean(imputedDaily$steps)
```

## Median value
```{r}
median(imputedDaily$steps)
```

Given my algorithm for filling in missing values it was obvious that mean / median values will change :) Surprisingly it was only a tiny bit.

Total daily values obvsiously grew:
```{r}
(sum(imputed$steps)-sum(data$steps, na.rm=TRUE)) / sum(data$steps, na.rm=TRUE)
```

# Are there differences in activity patterns between weekdays and weekends?
```{r}
imputed$day_type <- as.factor(ifelse(weekdays(imputed$date) %in% c("Saturday", "Sunday"), "weekend", "weekday"))
aggregated <- aggregate(steps ~ interval + day_type, imputed, FUN=mean)
xyplot(aggregated$steps ~ aggregated$interval | aggregated$day_type, type = "l")
```